---
title: "covid 19"
author: "Marium Khanum"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex?
$H_0$: There is no relationship between sex and the amount of people dying from Covid.  
$H_A$: There is a relationship between sex and the amount of people dying from Covid. 

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="Percentage Bar Chart")
```

As shown in the graph, the percentage of females dying from the Covid is less than that of males. The percentage of males with the Covid is less than females however the rates of males dying is more than females.  

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The results show that out of the whole population, the females alive is 56% whereas the males alive is 44%.  
Of all the females, 98.7% females are alive and 1.4% of them are dead.  
Of all the males, 97% males are alive and 3% of them are dead which is almost twice more than the percent of females that died from the virus.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of dying if you are a male is 44/1411.  
The odds of dying if you are a female is 25/1831.  
The odds ratio is (44/1411)/(25/1831)= 2.3 to 1.  

## Conclusions
The results shows that the p value is 0.001 which is less than 0.05 therefore we reject the null hypothesis. There is a relationship between sex and the amount of people that died due to Covid. 
